Large-Scale Marine Protected Areas in the World's Largest Tuna Fishery

### 2019-03-04 docs/old_ms/PIPA_Palau.tex
Updated MS

% Please add here a significance statement to explain the relevance of your work
\significancestatement{The oceans are becoming increasingly crowded, with different activities competing for space, but little research has been done to understand how activities displace each other. Marine Protected Areas are inherently spatial and have the sole objective of conserving bounded waters through the displacement of fishing effort. Our work shows that displacement of fishing effort occurs, but that it comes at a cost to countries that would have otherwise been able to charge a price for vessels to fish in the now-protected areas. With global conservation targets seeking to protect 10\% of the world’s ocean by 2020 and the expansion of offshore aquaculture, it is important that the spatial displacement of economic activities is fully considered.}

@@

\begin{abstract}
The Parties to the Nauru Agreement (PNA) is a system for managing the world’s largest skipjack tuna fishery, covering an area of 14.6 million km\textsuperscript{2} in the Pacific Ocean. The fishery in the region operates under a Vessel-Day Scheme (VDS), which sells access rights that allow vessels to fish in PNA waters. In 2015, PNA member Kiribati implemented one of the world's largest conservation areas: The Phoenix Islands Protected Area (PIPA, 397,447 km\textsuperscript{2}), effectively excluding all tuna purse seining activities. Such an intervention is likely to have effects on vessel spatial distribution and behavior, as well as induce costs (due to the reduction in fishable area) to Kiribati. We use identification of fishing activity via Automatic Identification Systems and causal inference techniques to evaluate the effect of PIPA on vessel distribution, behavior, and costs to Kiribati and the PNA. We find a crowding effect within PNA waters after the implementation of the protected area. Vessels continue to fish with similar intensity after the implementation. In the first year of the closure, there is no drop in total fishing effort within Kiribati's EEZ and a reported increase in revenue from access rights sold. However, from 2016 onwards there is a noticeable drop in fishing effort within Kiribati and a reported drop in VDS revenue. At the same time, fishing effort increases in other parts of the PNA. This redistribution of fishing effort eventually results in a reduction of 2,310 vessel-days in Kiribati, which represents a loss of \$27.7 million USD; similar estimates are obtained when looking at country-level license fees revenues directly (\$30.5 million USD). We use our results to inform predictions of the impacts of a proposed Large-Scale Marine Protected Area (LSMPA) in Palau (another PNA member) and estimate potential losses to range from \$2.5 to \$11 million annually. PNA members who indirectly benefit from MPAs should consider mechanisms that reward such conservation actions.
\end{abstract}

@@

\dropcap{H}umans are increasingly utilizing the oceans. Multiple ocean uses such as off-shore aquaculture, conservation, energy harvesting, deep-sea mining, and fisheries are likely to compete for space. As we move forward with blue growth, we must understand the potential effects of activities displacing each other and establish causal links between past management interventions and their outcomes \citep{burgess_2018}. One of the most notable spatial interventions is the creation of no-take Marine Protected Areas (MPAs), which seek to conserve the environment by eliminating fishing effort within their waters.

Global international goals aim to protect 10\% of the ocean environment by 2020. In an effort to meet this target, there has been a rapid increase in MPA coverage \citep{wood_2008,sala_2018}, largely driven by a small number of Large-Scale Marine Protected Areas (LSMPAs), some of which occur in the world's largest fishing grounds for tuna. While today a small number of LSMPAs make up at least 80\% of the managed areas in the ocean \citep{toonen_2013}, very little is known about their human dimensions and implications for fisheries \citep{gray_2017}. Furthermore, most of the research on LSMPAs has focused on their potential ecological benefits, but have left aside the economic implications. One issue of particular importance is that of the displacement or redistribution of fishing effort, which may influence the outcomes of a spatial closure and represent large opportunity costs \citep{smith_2003,smith_2010}.

The Phoenix Islands Protected Area (PIPA) in Kiribati is one of the most notable Large-Scale Marine Protected Areas. Implemented on January 1st of 2015, PIPA closed an area of 397,447 km\textsuperscript{2} to fishing and was implemented within an area that produces approximately 50\% of the world's tuna. Tuna purse seine fisheries in the region are collectively managed under a Vessel-Day Scheme (VDS) by nine countries commonly refered to as the Parties to the Nauru Agreement (PNA). Members include the Federated States of Micronesia, Kiribati, the Marshall Islands, Nauru, Palau, Papua New Guinea, the Solomon Islands, and Tuvalu; Tokelau joined the PNA group in 2012 and started selling access rights in 2013 (Figure \ref{fig:PNA_map}). The Nauru Agreement regulates access of foreign vessels (\emph{i.e.} those from non-PNA countries). Holding 80\% (14.6 million km\textsuperscript{2}) of historical skipjack tuna purse seine grounds within their Exclusive Economic Zones (EEZ), PNA countries have achieved greater bargaining power when providing fishing access to foreign fleets \cite{havice_2010}. The revenue from access fees may represent up to 50\% of government revenue for some of the members.

A spatial closure of such dimensions is likely to cause changes in spatial distribution and behavior of fishing vessels. For example, the anticipation of LSMPAs may lead to preemptive overfishing, which will likely erodes or delay the expected benefits of an intervention \cite{mcdermott_2018,hanich2018unraveling}. After a spatial closure, users can exhibit idiosyncratic responses driven by different incentives \cite{cabral_2017}. Moreover, this reduction in total fishing area within one country's EEZ is likely to result in selling less fishing licenses. While no studies have assessed the implications of PIPA, other PNA members have pledged the implementation of LSMPAs by 2020 (i.e. Palau). We evaluate the behavioral responses and spatial redistribution of the industrial tuna purse seine fleet resulting from the implementation of the Phoenix Islands Protected Area, and quantify its economic ramifications and impacts to Kiribati. We use the same data to hypothesize what might be the impacts of the proposed Palau National Marine Sanctuary. These are two of the largest protected areas on the planet and both are controled by PNA countries, where the largest tuna fisheries occur.

We use identification of fishing activity via Automatic Identification Systems (AIS) to track 313 tuna purse seine vessels that fished in PNA waters between 2012 and 2018. We continuously observe 92 vessels for the 2012 - 2018 period. Of these, 64 vessels fished within PIPA at least once prior to its implementation. The remaining 28 vessels never fished in PIPA waters. We refer to these groups as the treated and control groups. The group with the remaining 221 vessels contains vessels that were not observed before and after the implementation of PIPA, and we refer to these as ``other vessels''.

\begin{figure}
\centering
\includegraphics{img/LSMPAs_map.pdf}
\caption{\label{fig:LSMPAs_map}Large Scale Marine Protected Areas. The map shows all areas larger than 30,000 Km\textsuperscript{2}. The Phoenix Islands Protected Area is shown in red.}
\end{figure}

@@

We first inspect the crowding effects that may arise due to the net reduction in fishing area. We produced 1-degree rasters of monthly fishing effort for our treatment and control groups, and calculated two indices of spatial overlap between them: 1) the number of cells that had fishing activity from treated and control vessels for each month and 2) the correlation of presence/absence of fishing events between both groups over one month. We find that the two fleets significantly interact more with each other after the implementation of PIPA (Table \ref{tab:sp_corr} Fig. \ref{fig:sp_corr}). The number of cells with presence from both fleets and spatial correlations increase by a factor of four and three, respectively. This increase in crowding is likely to increase the encounter rates with other vessels, and reduce the efficiency of fishing operations. This might cause vessels to leave their current fishing grounds and re-optimize their spatial effort, leading to a subsequent decrease as the crowding measures return towards pre-implementation levels.

@@

The behavioral responses that vessels can have to a spatial closure may occurr in different ways. For example, displacement to new fishing grounds may represent a cost as fishers search the ocean to identify the most suitable fishing spots. This may result in increased fuel and labor costs. For every vessel in each group, we calculate eight key measures that could capture responses to spatial closures: daily fishing hours, daily non-fishing at-sea hours, the proportion of fishing to non-fishing hours at sea, daily distance traveled, daily mean distance from shore of fishing events (km), daily mean distance from port of fishing events (km), as well as monthly hours spent in PNA waters and Kiribati waters (Fig. \ref{fig:all_panels}). We leverage our Before-After-Control-Impact (BACI) design and implement a log-linear difference-in-differences analysis to evaluate how these measures change for treated vessels after implementation of PIPA, relative to the trends observed for control vessels (see the Methods section for our empirical specification).

We find no evidence of treated vessels fishing for more hours after PIPA implementation, and a slight decrease of fishing hours relative to total at-sea hours ($p < 0.01$; Table \ref{tab:main_DID}). Treated vessels traveled less, with fishing events occurring closer to shore and closer to port following PIPA implementation. These changes in distance from shore and port are likely caused by redistribution, as we observe that treated vessels fish 56.5\% and 39.9\% less in Kiribati and PNA waters, compared to the trend observed for control vessels ($p < 0.01$). We do not observe a statistically significant increase in fishing hours on the high seas.

The fact that the amount of time that vessels spent at sea (fishing and non-fishing) did not decrease, and that their time spent in Kiribati and PNA waters decreases suggests that these vessels have re-distributed elsewhere. This is consistent with the eventual decrease in crowding meassures.

>>>>

% Please add here a significance statement to explain the relevance of your work
\significancestatement{The oceans are becoming increasingly crowded, with different activities competing for space, but little research has been done to understand how activities displace each other. Marine Protected Areas are inherently spatial and have the sole objective of conserving bounded waters through the displacement of fishing effort. Our work shows that displacement of fishing effort occurs, but that it comes at a large cost to countries that would have otherwise been able to charge a price for vessels to fish in the now-protected areas. With global conservation targets seeking to protect 10\% of the world’s ocean by 2020 and the expansion of offshore aquaculture, it is important that the spatial displacement of economic activities is fully considered.}

@@

\begin{abstract}
The Parties to the Nauru Agreement (PNA) is a system for managing the world’s largest skipjack tuna fishery, covering an area of 14.6 million km\textsuperscript{2} in the Pacific Ocean. The fishery in the region operates under a vessel-day scheme (VDS), which sells access rights that allow vessels to fish in PNA waters. In 2015, PNA member Kiribati implemented one of the world's largest conservation areas: The Phoenix Island Protected Area (PIPA, 397,447 km\textsuperscript{2}), effectively excluding all tuna purse seining activities. Such an intervention is likely to have effects on vessel spatial distribution and behavior, as well as induce costs (due to the reduction in fishable area) to Kiribati. We use identification of fishing activity via Automatic Identification Systems and causal inference techniques to evaluate the effect of PIPA on vessel distribution, behavior, and costs to Kiribati and the PNA. We find a short-term crowding effect within PNA waters after the implementation of the protected area. Vessels continue to fish with similar intensity after the implementation. The eventual decrease in crowding within PNA waters but constant fishing intensity suggests that vessels redistribute to adjacent waters, including the high seas. This redistribution results in a reduction of 2,310 vessel-days in Kiribati, which represents a loss of \$27.7 million USD; similar estimates are obtained when looking at country-level license fees revenues directly (\$30.5 million USD). We use our results to inform predictions of the impacts of a proposed LSMPA in Palau (a PNA member) and estimate potential losses to range from \$2.5 to \$11 million annually. PNA members who decide to implement MPAs should consider mechanisms that reward such conservation actions.
\end{abstract}

@@

\dropcap{H}umans are increasingly utilizing the oceans. Multiple ocean uses such as off-shore aquaculture, conservation, energy harvesting, deep-sea mining, and fisheries are likely to compete for space. As we move forward with blue growth, we must understand the potential effects of activities displacing each other and etablish causal links between past management interventions and their outcomes \citep{burgess_2018}. One of the most notable spatial spatial interventions are no-take Marine Protected Areas, which seek to conserve the environment by eliminating fishing effort within their waters.

Global international goals aim to protect 10\% of the ocean environment by 2020. In an effort to meet this target, there has been a rapid increase in MPA coverage \citep{wood_2008,sala_2018}, largely driven by a small number of Large-Scale Marine Protected Areas (LSMPAs), some of which ocurr in the world's largest fishing grounds for tuna. While today a small number of LSMPAs make up at least 80\% of the managed areas in the ocean \citep{toonen_2013}, very little is known about their human dimensions and implication for fisheries \citep{gray_2017}. Furthermore, most of the research on LSMPAs has focused on their potential ecological benefits, but have left aside the economic implications. One issue of particular importance is that of the displacement or redistribution of fishing effort, which may influence the outcomes of a spatial closure and represent large opportunity costs \citep{smith_2003,smith_2010}.

The Phoenix Island Protected Area (PIPA) in Kiribati is one of the most notable Large-Scale Marine Protected Areas. Implemented in January 1st of 2015, PIPA has an extension of 397,447 km\textsuperscript{2} and was implemented within an area that produces 50\% of the World's tuna. Tuna purse seine fisheries in the region are collectively managed under a Vessel-day scheme (VDS) by nine countries commonly refered to as the Parties to the Nauru Agreement (PNA). Members include the Federated States of Micronesia, Kiribati, the Marshall Islands, Nauru, Palau, Papua New Guinea, the Solomon Islands, and Tuvalu. Tokelau joined the PNA group in 2012 and started selling access rights in 2013. The Nauru Agreement regulates access of foreign vessels (\emph{i.e.} those from non-PNA countries). Holding 80\% (14.6 million km\textsuperscript{2}) of historical skipjack tuna purse seine grounds within their Exclusive Economic Zones (EEZ), PNA countries have achieved greater bargaining power when providing fishing access to foreign fleets \cite{havice_2010}. The revenue from access fees may represent up to 50\% of government revenue for some of the members.

A spatial closure of such dimensions is likely to cause changes in spatial distribution and behavior of fishing vessels. For example, the anticipation of LSMPAs may lead to preemptive overfishing, which will likely erodes or delay the expected benefits of an intervention \cite{mcdermott_2018}. After a spatial closure, users can exhibit idiosyncratic responses driven by different incentives \cite{cabral_2017}. Moreover, this reduction in total fishing area within one country's EEZ is likely to result in selling less fishing licenses. While no studies have assessed the immplications of PIPA, other PNA members have pledged the implementation of LSMPAs by 2020. We evaluate the behavioral responses and spatial redistribution of the industrial tuna purse seine fleet resulting from the implementation of the Phoenix Islands Protected Area, and quantify its economic ramifications and impacts to Kiribati. We use the same data to hypothesize what might be the impacts of the proposed Palau National Marine Sanctuary. These are two of the largest protected areas on the planet and both are controled by PNA countries, where the largest tuna fisheries ocurr.

We use identification of fishing activity via Automatic Identification Systems (AIS) to track 313 tuna purse seine vessels that fished in PNA waters between 2012 and 2018. We continuously observe 92 vessels for the 2012 - 2018 period. Of these, 64 vessels fished within PIPA at least once prior to its implementation. The remaining 28 vessels never fished in PIPA these waters. We refer to these groups as treated and control groups. The group with the remaining 221 vessels contains vessels that were not observed before and after the implementation of PIPA, and we refer to these as ``other vessels''.

\begin{figure}
\centering
\includegraphics{img/LSMPAs_map.pdf}
\caption{\label{fig:LSMPAs_map}Large Scale Marine Protected Areas. The map shows all areas larger than 30,000 Km\textsuperscript{2}. The Phoenix Island Protected Area is shown in red.}
\end{figure}

@@

We first inspect the crowding effects that may arise due to the net reduction in fishing area. We produced 1-degree rasters of monthly fishing effort for our treatment and control groups, and calculated two indices of spatial overlap between thm: 1) the number of cells that had fishing activity from treated and control vessels for each month and 2) the correlation of presence/absence of fishing events between both groups over one month. We find that the two fleets significantly interact more with each other after the implementation of PIPA (Table \ref{tab:sp_corr} Fig. \ref{fig:sp_corr}). The number of vessels with presence from both fleets and spatial correlations increase by a factor of four and three, respectively. This increase in crowding is likely to increase the encounter rates within vessels, and reduce the efficiency of fishing opperations, resulting in a nusicance to vessels. This would cause vessels to leave curent fishing grounds, leading to the subsequent decrease as the crowding measures return to pre-implementation levels.

@@

The behavioral responses that vessels can have to a spatial closure may occurr in different ways. For example, displacement to new fishing grounds may represent a cost as fishers search the ocean to identify the most suitable fishing places. This may result in increased fuel and labor costs. For every vessel in each group, we calculate in eight key measures that could capture responses to spatial closures: daily fishing hours, daily non-fishing at-sea hours, the proportion of fishing to non-fishing hours at sea, daily distance traveled, daily mean distance from shore of fishing events (km), daily mean distance from port of fishing events (km), as well as monthly hours spent in PNA waters and Kiribati waters (Fig. \ref{fig:all_panels}). We leverage our BACI design and implement a log-linear difference-in-differences analyses to evaluate how these measures change for treated vessels after implementation of PIPA, relative to the trends observed for control vessels (See Methods section for our empirical specification).

We find no evidence of treated vessels fishing for more hours after PIPA implementation, and a slight increase of 1\% more time at sea ($p < 0.05$; Table \ref{tab:main_DID}). Treated vessels traveled 20\% less, with fishing events ocurring 23.8\% closer to shore and 5\% closer to port. These changes in distance from shore and port are likely caused by redistribution, as we observe that treated vessels fish 58\% and 45\% less in PNA and Kiribati waters, compared to the trend observed for control vessels ($p < 0.01$). The fact that the amount of time that vessels spent at sea (fishing and non-fishing) did not decrease, and that their time spent in PNA and Kiribati waters decreases suggests that these vessels have re-distributed elsewhere. This is consistent with the eventual decrease in crowding meassures.

### 2019-03-04 docs/tab/sp_corr.tex
Updated MS

\begin{table}[!htbp] \centering 
  \caption{\label{tab:sp_corr}Coefficient estimates for a third-polinomial fit to the measures of crowding. The first column shows coefficients for the number of cells with treated and control vessels during the same month. The second column shows coefficients for the spatial correlation for presence / absence of treated and control vessels. The explanatory variable is the number of months before implementation of PIPA. Numbers in parentheses are heteroskedastic-robust standard errors.} 
  \label{} 
\footnotesize 
\begin{tabular}{@{\extracolsep{1pt}}lcc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
\\[-1.8ex] & (1) & (2)\\ 
\hline \\[-1.8ex] 
 Constant & 59.999$^{***}$ (15.581) & 0.370$^{***}$ (0.062) \\ 
  Months & 3.014$^{***}$ (0.967) & 0.009$^{**}$ (0.003) \\ 
  Months $^2$ & 0.010 (0.027) & $-$0.0001 (0.0001) \\ 
  Months $^3$ & $-$0.002$^{***}$ (0.001) & $-$0.00001$^{***}$ (0.00000) \\ 
  Months $^4$ & 0.00000 (0.00002) & 0.00000 (0.00000) \\ 
  sate1 & 22.342 (19.528) & 0.057 (0.076) \\ 
  sate2 & 5.413 (18.860) & $-$0.015 (0.035) \\ 
 \hline \\[-1.8ex] 
Month FE & Yes & Yes \\ 
Flag FE & Yes & Yes \\ 
Observations & 83 & 84 \\ 
R$^{2}$ & 0.791 & 0.709 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

>>>>

\begin{table}[!htbp] \centering 
  \caption{\label{tab:sp_corr}Coefficient estimates for a third-polinomial fit to the measures of crowding. The first column shows coefficients for the number of cells with treated and control vessels during the same month. The second column shows coefficients for the spatial correlation for presence / absence of treated and control vessels. The explanatory variable is the number of months before implementation of PIPA. Numbers in parentheses are heteroskedastic-robust standard errors.} 
  \label{} 
\footnotesize 
\begin{tabular}{@{\extracolsep{1pt}}lcc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
\\[-1.8ex] & (1) & (2)\\ 
\hline \\[-1.8ex] 
 Constant & 59.999$^{***}$ (15.581) & 0.370$^{***}$ (0.062) \\ 
  $M$ & 3.014$^{***}$ (0.967) & 0.009$^{**}$ (0.003) \\ 
  $M^2$ & 0.010 (0.027) & $-$0.0001 (0.0001) \\ 
  $M^3$ & $-$0.002$^{***}$ (0.001) & $-$0.00001$^{***}$ (0.00000) \\ 
  $M^4$ & 0.00000 (0.00002) & 0.00000 (0.00000) \\ 
  $\sigma_1$ & 22.342 (19.528) & 0.057 (0.076) \\ 
  $\sigma_2$ & 5.413 (18.860) & $-$0.015 (0.035) \\ 
 \hline \\[-1.8ex] 
Month FE & Yes & Yes \\ 
Flag FE & Yes & Yes \\ 
Observations & 83 & 84 \\ 
R$^{2}$ & 0.791 & 0.709 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}
